---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data


Read in the data from the activity file

```{r}
data <- read.csv("activity.csv", header = TRUE, sep = ',', colClasses = c("numeric", "character","integer"))
```


Load in the libraries being used.

```{r}
library(dplyr)
library(lubridate)
library(ggplot2)
```

Change date into dateformat using as.Date
```{r}
data$date <- as.Date(data$date)
```


```{r}
head(data)
```


## What is mean total number of steps taken per day?

 1) Calculate total steps taken in a day:
 
Aggregate the number of steps in a day by date, using group_by
```{r}
steps <- data %>%
  group_by(date) %>%
  summarize(steps = sum(steps,na.rm=TRUE)) 
steps

```

2) Make a histrogram of total steps perday



```{r}
ggplot(steps, aes(x = steps)) + geom_histogram(fill = "blue", binwidth = 1000) + labs(title = "Steps per day", x = "Steps per day", y = "Frequency")
```

 3) Calculate the mean and median of the total number of steps taken per day
```{r}

 
mean_steps <- mean(steps$steps, na.rm = TRUE)
median_steps <- median(steps$steps, na.rm = TRUE)

mean_steps

median_steps
```


## What is the average daily activity pattern?

a) Calculate the average number of steps taken in each 5-minute interval per day 

```{r}
intv <- data  %>%
  group_by(interval) %>%
  summarize(steps = mean(steps,na.rm=TRUE))
```

Plot

```{r}
ggplot(intv, aes(x=interval, y=steps)) + geom_line(color = "blue")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?.

 
```{r}
intv[which.max(intv$steps),]
```



## Imputing missing values


Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs).
```{r}
sum(is.na(data$steps))

```

Create a new dataset that is equal to the original dataset but with the missing data filled in by using mean steps in an interval.

```{r}
imputed.data<- data %>%
  group_by(interval)  %>%
  mutate(steps= ifelse(is.na(steps), mean(steps, na.rm=TRUE),steps))

head(imputed.data)
```


Check to see if there any missing values. There are none.
```{r}
sum(is.na(imputed.data$steps))

```

Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.
```{r}
steps_imputed<- imputed.data %>%
  filter(!is.na(steps)) %>%
  group_by(date) %>% summarize(steps = sum(steps)) 

ggplot(steps_imputed, aes(x = steps)) +
  geom_histogram(fill = "blue", binwidth = 1000) +
  labs(title = "Histogram of Steps per day, including missing values", x = "Steps per day", y = "Frequency")

```

Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r}
mean_steps_imputed <- mean(steps_imputed$steps, na.rm = TRUE)
mean_steps_imputed


median_steps_imputed <- median(steps_imputed$steps, na.rm = TRUE)
median_steps_imputed
```
The impact of imputing missing data with the average number of steps in the same 5-min interval is that both the mean and the median are equal to the same value: 10766.

## Are there differences in activity patterns between weekdays and weekends?


Use dplyr and mutate to create a new column, weektype, and apply whether the day is weekend or weekday:

```{r}
imputed.data2 <- mutate(imputed.data, weektype = weekdays(date)) %>% mutate(weektype=ifelse(weektype=="Sunday"|weektype=="Sunday",'weekend','weekday'))
imputed.data2$weektype <- as.factor(imputed.data2$weektype)
head(imputed.data2)
```


Calculate the average steps in the 5-minute interval and use ggplot for making the time series of the 5-minute interval for weekday and weekend, and compare the average steps:
```{r}
interval_full <- imputed.data2 %>%
  group_by(interval, weektype) %>%
  summarise(steps = mean(steps))
s <- ggplot(interval_full, aes(x=interval, y=steps, color = weektype)) +
  geom_line() +
  facet_wrap(~weektype, ncol = 1, nrow=2)
print(s)
```

