---
title: "peer assignment"
author: "Nishant Sahni"
date: "November 13, 2015"
output: html_document
---

# Preprocessing files and the dataset

Read in the data from the activity file

```{r}
data <- read.csv("activity.csv", header = TRUE, sep = ',', colClasses = c("numeric", "character","integer"))
```


Load in the libraries being used.

```{r}
library(dplyr)
library(lubridate)
library(ggplot2)
```

Change date into dateformat using as.Date
```{r}
data$date <- as.Date(data$date)
```


```{r}
head(data)
```


## 1) Calculate total steps taken in a day:
Aggregate the number of steps in a day by date, using group_by
```{r}
steps <- data %>%
  group_by(date) %>%
  summarize(steps = sum(steps,na.rm=TRUE)) 
steps

```

## 2) Make a histrogram of total steps perday



```{r}
ggplot(steps, aes(x = steps)) + geom_histogram(fill = "blue", binwidth = 1000) + labs(title = "Steps per day", x = "Steps per day", y = "Frequency")
```

## 3) Calculate the mean and median of the total number of steps taken per day
```{r}

 
mean_steps <- mean(steps$steps, na.rm = TRUE)
median_steps <- median(steps$steps, na.rm = TRUE)

mean_steps

median_steps
```



## What is the average daily activity pattern?

## 1) Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis).

a) Calculate the average number of steps taken in each 5-minute interval per day 

```{r}
intv <- data  %>%
  group_by(interval) %>%
  summarize(steps = mean(steps,na.rm=TRUE))
```


```{r}
ggplot(intv, aes(x=interval, y=steps)) + geom_line(color = "blue")
```

## Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?.

 
```{r}
intv[which.max(intv$steps),]
```


#Imputing missing values


Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs).
```{r}
sum(is.na(data$steps))

```

## Create a new dataset that is equal to the original dataset but with the missing data filled in by using mean steps in an interval.

```{r}
imputed.data<- data %>%
  group_by(interval)  %>%
  mutate(steps= ifelse(is.na(steps), mean(steps, na.rm=TRUE),steps))

head(imputed.data)
```


Check to see if there any missing values. There are none.
```{r}
sum(is.na(imputed.data$steps))

```

